### 2023-08-11 scripts/Publilia-03-survivorship.Rmd

---
title: "Publilia-03-survivorship.Rmd"
author: "micah fletcher"
date: "8/4/2023"
output: html_document
---

## Required libraries and functions
```{r}
library(tidyverse)
library(lubridate)
library(ggpubr)
library(ggh4x)
library(rstatix)
library(emmeans)
library(effectsize)
library(gt)
library(gtsummary)
theme_gtsummary_compact()
library(flextable)
library(multcompView)
library(here)

theme_publilia <- function(){
  theme_bw() +
    theme(text = element_text(family = "Arial"),
          axis.text = element_text(size = 12), 
          axis.title = element_text(size = 14),
          axis.line.x = element_line(color="black"), 
          axis.line.y = element_line(color="black"),
          panel.border = element_blank(),
          panel.grid.major.x = element_blank(),                                          
          panel.grid.minor.x = element_blank(),
          panel.grid.minor.y = element_blank(),
          panel.grid.major.y = element_blank(),  
          plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), units = , "cm"),
          plot.title = element_text(size = 18, vjust = 1, hjust = 0.5, face = "bold"),
          legend.text = element_text(size = 10),          
          legend.title = element_blank(),                              
          # legend.position = c(0.90, 0.90), 
          legend.key = element_blank(),
          legend.background = element_rect(color = "black", 
                                           fill = "transparent", 
                                           size = 2, linetype = "blank"))
}
```

# 0 - Data import and custom functions

## Data for regressions
```{r}
surv_reg_data <- read_rds(here("saved_rds/surv_reg_data.rds"))
```

## Custom functions

### Model fitting and confidence intervals and plotting functions
```{r}
### fit survivorship model
fit_surv <- function(data_in, survNom, survDen, tenure, 
                     treatment, vis, interaction=c("full","pw")){
  data <- eval(substitute(data_in %>% filter(!is.na(tenure))))

  mai.b <- eval(substitute(glm(survNom/survDen ~ 
            tenure*treatment,
          data = data, weights = survDen, family = binomial)))
  print("Interaction, Binomial: ")
  print(capture.output(summary(mai.b))[-(6:which(grepl("Deviance Residuals: ", capture.output(summary(mai.b))))-1)])

  mai.qb <- eval(substitute(glm(survNom/survDen ~ 
            tenure*treatment,
          data = data, weights = survDen, family = quasibinomial)))
  print("Interaction, Quasibinomial: ")
  print(capture.output(summary(mai.qb))[-(6:which(grepl("Deviance Residuals: ", capture.output(summary(mai.qb))))-1)])

  ma.qb <- eval(substitute(glm(survNom/survDen ~ 
            tenure+treatment,
          data = data, weights = survDen, family = quasibinomial)))
  print("No Interaction, Quasibinomial: ")
  print(capture.output(summary(mai.qb))[-(6:which(grepl("Deviance Residuals: ", capture.output(summary(mai.qb))))-1)])

  mai.models <- list(mai.b = mai.b, mai.qb = mai.qb, ma.qb = ma.qb)
  return(mai.models)
}

# Example
# test_fits <- fit_surv(data_in=hatch.data, tenure=des, treatment=ants,
#                  interaction="pw", survNom=n_hatched, survDen=med_t5eggs)

### fit survivorship model
fit_surv_quad <- function(data_in, survNom, survDen, tenure, 
                     treatment){
  data <- eval(substitute(data_in %>% filter(!is.na(tenure))))

  mai.b <- eval(substitute(glm(survNom/survDen ~
            tenure*treatment + (I(as.numeric(tenure)^2))*treatment,
          data = data, weights = survDen, family = binomial)))
  print("Interaction, Binomial: ")
  print(capture.output(summary(mai.b))[-(6:which(grepl("Deviance Residuals: ", capture.output(summary(mai.b))))-1)])

  mai.qb <- eval(substitute(glm(survNom/survDen ~
            tenure*treatment + (I(as.numeric(tenure)^2))*treatment,
          data = data, weights = survDen, family = quasibinomial)))
  print("Interaction, Quasibinomial: ")
  print(capture.output(summary(mai.qb))[-(6:which(grepl("Deviance Residuals: ", capture.output(summary(mai.qb))))-1)])

  ma.qb <- eval(substitute(glm(survNom/survDen ~
            tenure+treatment + (I(as.numeric(tenure)^2))*treatment,
          data = data, weights = survDen, family = quasibinomial)))
  print("No Interaction, Quasibinomial: ")
  print(capture.output(summary(ma.qb))[-(6:which(grepl("Deviance Residuals: ", capture.output(summary(ma.qb))))-1)])

    mai.models <- list(mai.b = mai.b, mai.qb = mai.qb, ma.qb = ma.qb)
  return(mai.models)
}

# Example
# test_fits <- fit_surv(data_in=hatch.data, tenure=des, treatment=ants,
#                  survNom=n_hatched, survDen=med_t5eggs)

### Create confidence intervals in the link scale for plotting, and backtransform for plot
# Based on: https://fromthebottomoftheheap.net/2018/12/10/confidence-intervals-for-glms/
ci_for_plot <- function(data_in, survNom, survDen, 
                        tenure, treatment, model){

  data <- eval(substitute(
    data_in %>% filter(!is.na(tenure)) %>%
      select(location, tenure, treatment, survDen, survNom) %>%
      add_column(fit = predict(model, newdata = ., type = 'response'))
    ))

  ## grad the inverse link function
  ilink <- family(model)$linkinv

  ## add fit and se.fit on the **link** scale
  data_fit_link <- eval(substitute(
    data %>% add_column(
  setNames(as_tibble(predict(model, data, se.fit = TRUE)[1:2]),
           c('fit_link','se_link'))) %>%
## create the interval and backtransform
  mutate(fit_resp  = ilink(fit_link),
         conf_upr = ilink(fit_link + (2 * se_link)),
         conf_lwr = ilink(fit_link - (2 * se_link)))
    ))

  return(data_fit_link)
}

# Example
# test_fits.pw <- fit_surv(data_in=hatch.data, tenure=des, treatment=ants, 
#                  interaction="pw", survNom=n_hatched, survDen=med_t5eggs)
# 
# test_ci_data <- ci_for_plot(data_in=hatch.data, tenure=des, treatment=ants, 
#                  survNom=n_hatched, survDen=med_t5eggs,
#                  model = test_fits.pw$mai.qb)

### Plot model results with confidence intervals
plot_ci <- function(ci_data, survNom, survDen, treatment, tenure){
  (eval(substitute(plot <- ggplot(data = ci_data, 
       aes(x = tenure, y = fit, 
           color = treatment, fill = treatment)) +
    theme_publilia() +
    geom_point(aes(y = survNom/survDen), alpha = 0.25, size = 4, shape=16, key_glyph=draw_key_rect) + 
    geom_line(aes(y = fit), size = 1, key_glyph=draw_key_rect) +
    geom_ribbon(aes(ymin = conf_upr, ymax = conf_lwr),
                    alpha = 0.5, color = NA, key_glyph=draw_key_rect))))
#
}

# Example
# test_fits.pw <- fit_surv(data_in=hatch.data, tenure=des, treatment=ants,
#                  interaction="pw", survNom=n_hatched, survDen=med_t5eggs)
# 
# test_ci_data <- ci_for_plot(data_in=hatch.data, tenure=des, treatment=ants,
#                  survNom=n_hatched, survDen=med_t5eggs,
#                  model = test_fits.pw$mai.qb)
# 
# (p <- plot_ci(ci_data = test_ci_data, tenure = des, treatment=ants,
#              survNom = n_hatched, survDen = med_t5eggs) +
#     xlab("Latest desertion date relative to hatch") +
#     ylab("Hatching success") +
#     scale_color_manual(values = c("brown1", "black")) +
#     scale_fill_manual(values = c("brown1", "black")))
# summary(test_fits.pw$mai.qb)
```

# Supp 1 -`desL * ants_treatment * desL^2`
## Figures
```{r}
# hatching success
des_hatch.desL.ant_treatment.quad.fits <- fit_surv_quad(data_in=surv_reg_data, 
                              survNom=n_hatched, survDen=med_t5eggs,
                              tenure=des, treatment=ant_treatment)

des_hatch.desL.ant_treatment.quad.qb.ci <- 
  ci_for_plot(data_in=surv_reg_data, 
              tenure=des, treatment=ant_treatment,
              survNom=n_hatched, survDen=med_t5eggs,
              model = des_hatch.desL.ant_treatment.quad.fits$mai.qb)

des_hatch.desL.ant_treatment.quad.qb.p <- plot_ci(ci_data=des_hatch.desL.ant_treatment.quad.qb.ci, 
                                  tenure=des, treatment=ant_treatment,
             survNom = n_hatched, survDen = med_t5eggs) +
    xlab("Latest desertion date from leaf\nrelative to hatch") +
    ylab("Hatching success\n(nymphs/eggs)") + 
    scale_color_manual(values = c("brown1", "black")) +
    scale_fill_manual(values = c("brown1", "black"))

# nymph survivorship
des_Nsurv.desL.ant_treatment.quad.fits <- fit_surv_quad(data_in=surv_reg_data, 
                              survNom=n_survived, survDen=n_hatched,
                              tenure=des, treatment=ant_treatment)

des_Nsurv.desL.ant_treatment.quad.qb.ci <- 
  ci_for_plot(data_in=surv_reg_data, 
              tenure=des, treatment=ant_treatment,
              survNom=n_survived, survDen=n_hatched,
              model = des_Nsurv.desL.ant_treatment.quad.fits$mai.qb)

des_Nsurv.desL.ant_treatment.quad.qb.p <- plot_ci(ci_data=des_Nsurv.desL.ant_treatment.quad.qb.ci, 
                                  tenure=des, treatment=ant_treatment,
             survNom = n_survived, survDen = n_hatched) +
    xlab("Latest desertion date from leaf\nrelative to hatch") +
    ylab("Survivorship of nymphs\n(5th instars/nymphs)") + 
    scale_color_manual(values = c("brown1", "black")) +
    scale_fill_manual(values = c("brown1", "black"))

# egg survivorship
des_Esurv.desL.ant_treatment.quad.fits <- fit_surv_quad(data_in=surv_reg_data, 
                              survNom=n_survived, survDen=med_t5eggs,
                              tenure=des, treatment=ant_treatment)

des_Esurv.desL.ant_treatment.quad.qb.ci <- 
  ci_for_plot(data_in=surv_reg_data, 
              tenure=des, treatment=ant_treatment,
              survNom=n_survived, survDen=med_t5eggs,
              model = des_Esurv.desL.ant_treatment.quad.fits$mai.qb)

des_Esurv.desL.ant_treatment.quad.qb.p <- plot_ci(ci_data=des_Esurv.desL.ant_treatment.quad.qb.ci, 
                                  tenure=des, treatment=ant_treatment,
             survNom = n_survived, survDen = med_t5eggs) +
    xlab("Latest desertion date from leaf\nrelative to hatch") +
    ylab("Survivorship of eggs\n(5th instars/eggs)") + 
    scale_color_manual(values = c("brown1", "black"), labels = c("Control", "Ant Exclusion")) +
    scale_fill_manual(values = c("brown1", "black"), labels = c("Control", "Ant Exclusion"))

(desL.ant_treatment.quad.qb.p <- egg::ggarrange(des_hatch.desL.ant_treatment.quad.qb.p+
                 theme(legend.position="none"),
               des_Nsurv.desL.ant_treatment.quad.qb.p+
                 theme(legend.position="none"),
               des_Esurv.desL.ant_treatment.quad.qb.p+
                 theme(legend.position=c(0.8,0.95))+ylim(0,1),
               labels=c("A","B","C"),
               ncol=3))

ggsave("desL.ant_treatment.quad.final.png",
       desL.ant_treatment.quad.qb.p,
       device="png", width = 300, height = 110, units = "mm", dpi = "print")
```

## Tables
```{r}
(surv_reg_data.tbl_summary.desL.ant_treatment <- surv_reg_data %>%
  mutate(ants = ant_treatment,
         females = ifelse(females_removed, "Females removed", "Control"),
         `leaf desert date rel. hatch` = des,
         `visited early` = vis,
         `n eggs` = med_t5eggs,
         `n hatched` = n_hatched,
         `n survived` = n_survived) %>%
  select(ants, females, `leaf desert date rel. hatch`, `visited early`, 
         `n eggs`, `n hatched`, `n survived`) %>%
  group_by(females, ants) %>%
  tbl_summary(by = ants) %>%
  bold_labels() %>% italicize_levels() %>%
  add_n())

surv_reg_data.tbl_summary.desL.ant_treatment %>%
  as_flex_table() %>%
  flextable::save_as_docx(path="surv_reg_data.tbl_summary.desL.ant_treatment.docx")
```

```{r}
hatch.desL.ant_treatment.quad.qb.tbl <- 
  tbl_regression(des_hatch.desL.ant_treatment.quad.fits$mai.qb, exponentiate=T,
                 label=list(des ~ "Leaf desert date",
                            `I(as.numeric(des)^2)` ~ "(Leaf desert date)^2",
                            ant_treatment ~ "Ants")) %>% bold_p()
Nsurv.desL.ant_treatment.quad.qb.tbl <- 
  tbl_regression(des_Nsurv.desL.ant_treatment.quad.fits$mai.qb, exponentiate=T,
                 label=list(des ~ "Leaf desert date",
                            `I(as.numeric(des)^2)` ~ "(Leaf desert date)^2",
                            ant_treatment ~ "Ants")) %>% bold_p()
Esurv.desL.ant_treatment.quad.qb.tbl <- 
  tbl_regression(des_Esurv.desL.ant_treatment.quad.fits$mai.qb, exponentiate=T,
                 label=list(des ~ "Leaf desert date",
                            `I(as.numeric(des)^2)` ~ "(Leaf desert date)^2",
                            ant_treatment ~ "Ants")) %>% bold_p()

(desL.ant_treatment.quad.qb.tbl.merge <- 
  tbl_merge(
    tbls = list(hatch.desL.ant_treatment.quad.qb.tbl, 
                Nsurv.desL.ant_treatment.quad.qb.tbl, 
                Esurv.desL.ant_treatment.quad.qb.tbl),
    tab_spanner = c("**Hatching success**", 
                    "**Survivorship of nymphs**", 
                    "**Survivorship of eggs**")
  ) %>%
  modify_header(label = "**Variable**") %>%
  bold_labels() %>% italicize_levels())

desL.ant_treatment.quad.qb.tbl.merge %>%
  as_flex_table() %>%
  flextable::save_as_docx(path="surv.reg_summary.desL.ants_treatment.quad.docx")
```

# Main - `desL * ants_gt2 * desL^2`

## Tables
```{r}
(surv_reg_data.tbl_summary.desL.ants_gt2d <- surv_reg_data %>%
  mutate(ants = ifelse(ants_gt2d==" ants", "ants >2 days", "ants <=2 days"),
         females = ifelse(females_removed, "Females removed", "Control"),
         `leaf desert date rel. hatch` = des,
         `visited early` = vis,
         `n eggs` = med_t5eggs,
         `n hatched` = n_hatched,
         `n survived` = n_survived) %>%
  select(ants, females, `leaf desert date rel. hatch`, `visited early`, 
         `n eggs`, `n hatched`, `n survived`) %>%
  group_by(females, ants) %>%
  tbl_summary(by = ants) %>%
  bold_labels() %>% italicize_levels() %>%
  add_n())

surv_reg_data.tbl_summary.desL.ants_gt2d %>%
  as_flex_table() %>%
  flextable::save_as_docx(path="surv_reg_data.tbl_summary.desL.ants_gt2d.docx")
```

```{r}
hatch.desL.ants_gt2d.quad.qb.tbl <- 
  tbl_regression(des_hatch.desL.ants_gt2d.quad.fits$mai.qb, exponentiate=T,
                 label=list(des ~ "Leaf desert date",
                            `I(as.numeric(des)^2)` ~ "(Leaf desert date)^2",
                            ants_gt2d ~ "Ants >2 days")) %>% bold_p()
Nsurv.desL.ants_gt2d.quad.qb.tbl <- 
  tbl_regression(des_Nsurv.desL.ants_gt2d.quad.fits$mai.qb, exponentiate=T,
                 label=list(des ~ "Leaf desert date",
                            `I(as.numeric(des)^2)` ~ "(Leaf desert date)^2",
                            ants_gt2d ~ "Ants >2 days")) %>% bold_p()
Esurv.desL.ants_gt2d.quad.qb.tbl <- 
  tbl_regression(des_Esurv.desL.ants_gt2d.quad.fits$mai.qb, exponentiate=T,
                 label=list(des ~ "Leaf desert date",
                            `I(as.numeric(des)^2)` ~ "(Leaf desert date)^2",
                            ants_gt2d ~ "Ants >2 days")) %>% bold_p()

(desL.ants_gt2d.quad.qb.tbl.merge <- 
  tbl_merge(
    tbls = list(hatch.desL.ants_gt2d.quad.qb.tbl, 
                Nsurv.desL.ants_gt2d.quad.qb.tbl, 
                Esurv.desL.ants_gt2d.quad.qb.tbl),
    tab_spanner = c("**Hatching success**", 
                    "**Survivorship of nymphs**", 
                    "**Survivorship of eggs**")
  ) %>%
  modify_header(label = "**Variable**") %>%
  bold_labels() %>% italicize_levels())

desL.ants_gt2d.quad.qb.tbl.merge %>%
  as_flex_table() %>%
  flextable::save_as_docx(path="surv.reg_summary.desL.ants_gt2d.quad.docx")
```

## Figures
```{r}
# hatching success
des_hatch.desL.ants_gt2d.quad.fits <- fit_surv_quad(data_in=surv_reg_data, 
                              survNom=n_hatched, survDen=med_t5eggs,
                              tenure=des, treatment=ants_gt2d)

des_hatch.desL.ants_gt2d.quad.qb.ci <- 
  ci_for_plot(data_in=surv_reg_data, 
              tenure=des, treatment=ants_gt2d,
              survNom=n_hatched, survDen=med_t5eggs,
              model = des_hatch.desL.ants_gt2d.quad.fits$mai.qb)

des_hatch.desL.ants_gt2d.quad.qb.p <- plot_ci(ci_data=des_hatch.desL.ants_gt2d.quad.qb.ci, 
                                  tenure=des, treatment=ants_gt2d,
             survNom = n_hatched, survDen = med_t5eggs) +
    xlab("Latest desertion date from leaf\nrelative to hatch") +
    ylab("Hatching success\n(nymphs/eggs)") + 
    scale_color_manual(values = c("brown1", "black")) +
    scale_fill_manual(values = c("brown1", "black"))

# nymph survivorship
des_Nsurv.desL.ants_gt2d.quad.fits <- fit_surv_quad(data_in=surv_reg_data, 
                              survNom=n_survived, survDen=n_hatched,
                              tenure=des, treatment=ants_gt2d)

des_Nsurv.desL.ants_gt2d.quad.qb.ci <- 
  ci_for_plot(data_in=surv_reg_data, 
              tenure=des, treatment=ants_gt2d,
              survNom=n_survived, survDen=n_hatched,
              model = des_Nsurv.desL.ants_gt2d.quad.fits$mai.qb)

des_Nsurv.desL.ants_gt2d.quad.qb.p <- plot_ci(ci_data=des_Nsurv.desL.ants_gt2d.quad.qb.ci, 
                                  tenure=des, treatment=ants_gt2d,
             survNom = n_survived, survDen = n_hatched) +
    xlab("Latest desertion date from leaf\nrelative to hatch") +
    ylab("Survivorship of nymphs\n(5th instars/nymphs)") + 
    scale_color_manual(values = c("brown1", "black")) +
    scale_fill_manual(values = c("brown1", "black"))

# egg survivorship
des_Esurv.desL.ants_gt2d.quad.fits <- fit_surv_quad(data_in=surv_reg_data, 
                              survNom=n_survived, survDen=med_t5eggs,
                              tenure=des, treatment=ants_gt2d)

des_Esurv.desL.ants_gt2d.quad.qb.ci <- 
  ci_for_plot(data_in=surv_reg_data, 
              tenure=des, treatment=ants_gt2d,
              survNom=n_survived, survDen=med_t5eggs,
              model = des_Esurv.desL.ants_gt2d.quad.fits$mai.qb)

des_Esurv.desL.ants_gt2d.quad.qb.p <- plot_ci(ci_data=des_Esurv.desL.ants_gt2d.quad.qb.ci, 
                                  tenure=des, treatment=ants_gt2d,
             survNom = n_survived, survDen = med_t5eggs) +
    xlab("Latest desertion date from leaf\nrelative to hatch") +
    ylab("Survivorship of eggs\n(5th instars/eggs)") + 
    scale_color_manual(values = c("brown1", "black"), labels = c("ants > 2 days", "ants ≤ 2 days")) +
    scale_fill_manual(values = c("brown1", "black"), labels = c("ants > 2 days", "ants ≤ 2 days"))

(desL.ants_gt2d.quad.qb.p <- egg::ggarrange(des_hatch.desL.ants_gt2d.quad.qb.p+
                 theme(legend.position="none"),
               des_Nsurv.desL.ants_gt2d.quad.qb.p+
                 theme(legend.position="none"),
               des_Esurv.desL.ants_gt2d.quad.qb.p+
                 theme(legend.position=c(0.8,0.95))+ylim(0,1),
               labels=c("A","B","C"),
               ncol=3))

ggsave("desL.ants_gt2d.quad.final.png",
       desL.ants_gt2d.quad.qb.p,
       device="png", width = 300, height = 110, units = "mm", dpi = "print")
```

# Supp 2 -`desP * ants_gt2 * desP^2`
## Figures
```{r}
# hatching success
des_hatch.desP.ants_gt2d.quad.fits <- fit_surv_quad(data_in=surv_reg_data, 
                              survNom=n_hatched, survDen=med_t5eggs,
                              tenure=desP, treatment=ants_gt2d)

des_hatch.desP.ants_gt2d.quad.qb.ci <- 
  ci_for_plot(data_in=surv_reg_data, 
              tenure=desP, treatment=ants_gt2d,
              survNom=n_hatched, survDen=med_t5eggs,
              model = des_hatch.desP.ants_gt2d.quad.fits$mai.qb)

des_hatch.desP.ants_gt2d.quad.qb.p <- plot_ci(ci_data=des_hatch.desP.ants_gt2d.quad.qb.ci, 
                                  tenure=desP, treatment=ants_gt2d,
             survNom = n_hatched, survDen = med_t5eggs) +
    xlab("Latest desertion date from plant\nrelative to hatch") +
    ylab("Hatching success\n(nymphs/eggs)") + 
    scale_color_manual(values = c("brown1", "black")) +
    scale_fill_manual(values = c("brown1", "black"))

# nymph survivorship
des_Nsurv.desP.ants_gt2d.quad.fits <- fit_surv_quad(data_in=surv_reg_data, 
                              survNom=n_survived, survDen=n_hatched,
                              tenure=desP, treatment=ants_gt2d)

des_Nsurv.desP.ants_gt2d.quad.qb.ci <- 
  ci_for_plot(data_in=surv_reg_data, 
              tenure=desP, treatment=ants_gt2d,
              survNom=n_survived, survDen=n_hatched,
              model = des_Nsurv.desP.ants_gt2d.quad.fits$mai.qb)

des_Nsurv.desP.ants_gt2d.quad.qb.p <- plot_ci(ci_data=des_Nsurv.desP.ants_gt2d.quad.qb.ci, 
                                  tenure=desP, treatment=ants_gt2d,
             survNom = n_survived, survDen = n_hatched) +
    xlab("Latest desertion date from plant\nrelative to hatch") +
    ylab("Survivorship of nymphs\n(5th instars/nymphs)") + 
    scale_color_manual(values = c("brown1", "black")) +
    scale_fill_manual(values = c("brown1", "black"))

# egg survivorship
des_Esurv.desP.ants_gt2d.quad.fits <- fit_surv_quad(data_in=surv_reg_data, 
                              survNom=n_survived, survDen=med_t5eggs,
                              tenure=desP, treatment=ants_gt2d)

des_Esurv.desP.ants_gt2d.quad.qb.ci <- 
  ci_for_plot(data_in=surv_reg_data, 
              tenure=desP, treatment=ants_gt2d,
              survNom=n_survived, survDen=med_t5eggs,
              model = des_Esurv.desP.ants_gt2d.quad.fits$mai.qb)

des_Esurv.desP.ants_gt2d.quad.qb.p <- plot_ci(ci_data=des_Esurv.desP.ants_gt2d.quad.qb.ci, 
                                  tenure=desP, treatment=ants_gt2d,
             survNom = n_survived, survDen = med_t5eggs) +
    xlab("Latest desertion date from plant\nrelative to hatch") +
    ylab("Survivorship of eggs\n(5th instars/eggs)") + 
    scale_color_manual(values = c("brown1", "black"), labels = c("ants > 2 days", "ants ≤ 2 days")) +
    scale_fill_manual(values = c("brown1", "black"), labels = c("ants > 2 days", "ants ≤ 2 days"))

(desP.ants_gt2d.quad.qb.p <- egg::ggarrange(des_hatch.desP.ants_gt2d.quad.qb.p+
                 theme(legend.position="none"),
               des_Nsurv.desP.ants_gt2d.quad.qb.p+
                 theme(legend.position="none"),
               des_Esurv.desP.ants_gt2d.quad.qb.p+
                 theme(legend.position=c(0.8,0.95))+ylim(0,1),
               labels=c("A","B","C"),
               ncol=3))

ggsave("desP.ants_gt2d.quad.final.png",
       desP.ants_gt2d.quad.qb.p,
       device="png", width = 300, height = 110, units = "mm", dpi = "print")
```

## Tables
don't need to include the data summary in supp, since the only thing that changes is the median (IQR) for desertion
```{r}
(surv_reg_data.tbl_summary.desP.ants_gt2d <- surv_reg_data %>%
  mutate(ants = ifelse(ants_gt2d==" ants", "ants >2 days", "ants <=2 days"),
         females = ifelse(females_removed, "Females removed", "Control"),
         `plant desert date rel. hatch` = desP,
         `visited early` = vis,
         `n eggs` = med_t5eggs,
         `n hatched` = n_hatched,
         `n survived` = n_survived) %>%
  select(ants, females, `plant desert date rel. hatch`, `visited early`, 
         `n eggs`, `n hatched`, `n survived`) %>%
  group_by(females, ants) %>%
  tbl_summary(by = ants) %>%
  bold_labels() %>% italicize_levels() %>%
  add_n())

surv_reg_data.tbl_summary.desP.ants_gt2d %>%
  as_flex_table() %>%
  flextable::save_as_docx(path="surv_reg_data.tbl_summary.desP.ants_gt2d.docx")
```

```{r}
hatch.desP.ants_gt2d.quad.qb.tbl <- 
  tbl_regression(des_hatch.desP.ants_gt2d.quad.fits$mai.qb, exponentiate=T,
                 label=list(desP ~ "Leaf desert date",
                            `I(as.numeric(desP)^2)` ~ "(Leaf desert date)^2",
                            ants_gt2d ~ "Ants >2 days")) %>% bold_p()
Nsurv.desP.ants_gt2d.quad.qb.tbl <- 
  tbl_regression(des_Nsurv.desP.ants_gt2d.quad.fits$mai.qb, exponentiate=T,
                 label=list(desP ~ "Leaf desert date",
                            `I(as.numeric(desP)^2)` ~ "(Leaf desert date)^2",
                            ants_gt2d ~ "Ants >2 days")) %>% bold_p()
Esurv.desP.ants_gt2d.quad.qb.tbl <- 
  tbl_regression(des_Esurv.desP.ants_gt2d.quad.fits$mai.qb, exponentiate=T,
                 label=list(desP ~ "Leaf desert date",
                            `I(as.numeric(desP)^2)` ~ "(Leaf desert date)^2",
                            ants_gt2d ~ "Ants >2 days")) %>% bold_p()

(desP.ants_gt2d.quad.qb.tbl.merge <- 
  tbl_merge(
    tbls = list(hatch.desP.ants_gt2d.quad.qb.tbl, 
                Nsurv.desP.ants_gt2d.quad.qb.tbl, 
                Esurv.desP.ants_gt2d.quad.qb.tbl),
    tab_spanner = c("**Hatching success**", 
                    "**Survivorship of nymphs**", 
                    "**Survivorship of eggs**")
  ) %>%
  modify_header(label = "**Variable**") %>%
  bold_labels() %>% italicize_levels())

desP.ants_gt2d.quad.qb.tbl.merge %>%
  as_flex_table() %>%
  flextable::save_as_docx(path="surv.reg_summary.desP.ants_gt2d.quad.docx")
```

# Supp 3 -`desP * ant_s_gt2_treatment * desP^2`
## Figures
```{r}
# hatching success
des_hatch.desP.ant_treatment.quad.fits <- fit_surv_quad(data_in=surv_reg_data, 
                              survNom=n_hatched, survDen=med_t5eggs,
                              tenure=desP, treatment=ant_treatment)

des_hatch.desP.ant_treatment.quad.qb.ci <- 
  ci_for_plot(data_in=surv_reg_data, 
              tenure=desP, treatment=ant_treatment,
              survNom=n_hatched, survDen=med_t5eggs,
              model = des_hatch.desP.ant_treatment.quad.fits$mai.qb)

des_hatch.desP.ant_treatment.quad.qb.p <- plot_ci(ci_data=des_hatch.desP.ant_treatment.quad.qb.ci, 
                                  tenure=desP, treatment=ant_treatment,
             survNom = n_hatched, survDen = med_t5eggs) +
    xlab("Latest desertion date from plant\nrelative to hatch") +
    ylab("Hatching success\n(nymphs/eggs)") + 
    scale_color_manual(values = c("brown1", "black")) +
    scale_fill_manual(values = c("brown1", "black"))

# nymph survivorship
des_Nsurv.desP.ant_treatment.quad.fits <- fit_surv_quad(data_in=surv_reg_data, 
                              survNom=n_survived, survDen=n_hatched,
                              tenure=desP, treatment=ant_treatment)

des_Nsurv.desP.ant_treatment.quad.qb.ci <- 
  ci_for_plot(data_in=surv_reg_data, 
              tenure=desP, treatment=ant_treatment,
              survNom=n_survived, survDen=n_hatched,
              model = des_Nsurv.desP.ant_treatment.quad.fits$mai.qb)

des_Nsurv.desP.ant_treatment.quad.qb.p <- plot_ci(ci_data=des_Nsurv.desP.ant_treatment.quad.qb.ci, 
                                  tenure=desP, treatment=ant_treatment,
             survNom = n_survived, survDen = n_hatched) +
    xlab("Latest desertion date from plant\nrelative to hatch") +
    ylab("Survivorship of nymphs\n(5th instars/nymphs)") + 
    scale_color_manual(values = c("brown1", "black")) +
    scale_fill_manual(values = c("brown1", "black"))

# egg survivorship
des_Esurv.desP.ant_treatment.quad.fits <- fit_surv_quad(data_in=surv_reg_data, 
                              survNom=n_survived, survDen=med_t5eggs,
                              tenure=desP, treatment=ant_treatment)

des_Esurv.desP.ant_treatment.quad.qb.ci <- 
  ci_for_plot(data_in=surv_reg_data, 
              tenure=desP, treatment=ant_treatment,
              survNom=n_survived, survDen=med_t5eggs,
              model = des_Esurv.desP.ant_treatment.quad.fits$mai.qb)

des_Esurv.desP.ant_treatment.quad.qb.p <- plot_ci(ci_data=des_Esurv.desP.ant_treatment.quad.qb.ci, 
                                  tenure=desP, treatment=ant_treatment,
             survNom = n_survived, survDen = med_t5eggs) +
    xlab("Latest desertion date from plant\nrelative to hatch") +
    ylab("Survivorship of eggs\n(5th instars/eggs)") + 
    scale_color_manual(values = c("brown1", "black")) +
    scale_fill_manual(values = c("brown1", "black"))

(desP.ant_treatment.quad.qb.p <- egg::ggarrange(des_hatch.desP.ant_treatment.quad.qb.p+
                 theme(legend.position="none"),
               des_Nsurv.desP.ant_treatment.quad.qb.p+
                 theme(legend.position="none"),
               des_Esurv.desP.ant_treatment.quad.qb.p+
                 theme(legend.position=c(0.8,0.95))+ylim(0,1),
               labels=c("A","B","C"),
               ncol=3))

ggsave("desP.ant_treatment.quad.final.png",
       desP.ant_treatment.quad.qb.p,
       device="png", width = 300, height = 110, units = "mm", dpi = "print")
```

## Tables
don't need to include the data summary in supp, since the only thing that changes is the median (IQR) for desertion
```{r}
(surv_reg_data.tbl_summary.desP.ant_treatment <- surv_reg_data %>%
  mutate(ants = ifelse(ant_treatment==" ants", "ants >2 days", "ants <=2 days"),
         females = ifelse(females_removed, "Females removed", "Control"),
         `plant desert date rel. hatch` = desP,
         `visited early` = vis,
         `n eggs` = med_t5eggs,
         `n hatched` = n_hatched,
         `n survived` = n_survived) %>%
  select(ants, females, `plant desert date rel. hatch`, `visited early`, 
         `n eggs`, `n hatched`, `n survived`) %>%
  group_by(females, ants) %>%
  tbl_summary(by = ants) %>%
  bold_labels() %>% italicize_levels() %>%
  add_n())

surv_reg_data.tbl_summary.desP.ant_treatment %>%
  as_flex_table() %>%
  flextable::save_as_docx(path="surv_reg_data.tbl_summary.desP.ant_treatment.docx")
```

```{r}
hatch.desP.ant_treatment.quad.qb.tbl <- 
  tbl_regression(des_hatch.desP.ant_treatment.quad.fits$mai.qb, exponentiate=T,
                 label=list(desP ~ "Leaf desert date",
                            `I(as.numeric(desP)^2)` ~ "(Leaf desert date)^2",
                            ant_treatment ~ "Ants")) %>% bold_p()
Nsurv.desP.ant_treatment.quad.qb.tbl <- 
  tbl_regression(des_Nsurv.desP.ant_treatment.quad.fits$mai.qb, exponentiate=T,
                 label=list(desP ~ "Leaf desert date",
                            `I(as.numeric(desP)^2)` ~ "(Leaf desert date)^2",
                            ant_treatment ~ "Ants")) %>% bold_p()
Esurv.desP.ant_treatment.quad.qb.tbl <- 
  tbl_regression(des_Esurv.desP.ant_treatment.quad.fits$mai.qb, exponentiate=T,
                 label=list(desP ~ "Leaf desert date",
                            `I(as.numeric(desP)^2)` ~ "(Leaf desert date)^2",
                            ant_treatment ~ "Ants")) %>% bold_p()

(desP.ant_treatment.quad.qb.tbl.merge <- 
  tbl_merge(
    tbls = list(hatch.desP.ant_treatment.quad.qb.tbl, 
                Nsurv.desP.ant_treatment.quad.qb.tbl, 
                Esurv.desP.ant_treatment.quad.qb.tbl),
    tab_spanner = c("**Hatching success**", 
                    "**Survivorship of nymphs**", 
                    "**Survivorship of eggs**")
  ) %>%
  modify_header(label = "**Variable**") %>%
  bold_labels() %>% italicize_levels())

desP.ant_treatment.quad.qb.tbl.merge %>%
  as_flex_table() %>%
  flextable::save_as_docx(path="surv.reg_summary.desP.ant_treatment.quad.docx")
```

# qAICc
```{r}
# qAICc into separate tables
# qaicc
(qAICc_gt_tbl_hatch_quad <- tibble(
  Model = c("desert leaf + ant treatment + desert leaf*ant treatment + 
            desert leaf^2 + desert leaf^2*ant treatment",
            "desert plant + ant treatment + desert plant*ant treatment + 
            desert plant^2 + desert plant^2*ant treatment",
            "desert leaf + ants >2days + desert leaf*ants >2days + 
            desert leaf^2 + desert leaf^2*ants >2days",
            "desert plant + ants >2days + desert plant*ants >2days + 
            desert plant^2 + desert plant^2*ants >2days"),
  Response = c(rep("Hatching success",4)),
  qAICc = c(
    get_qaicc(des_hatch.desL.ant_treatment.quad.fits),
    get_qaicc(des_hatch.desP.ant_treatment.quad.fits),
    get_qaicc(des_hatch.desL.ants_gt2d.quad.fits),
    get_qaicc(des_hatch.desP.ants_gt2d.quad.fits))
  ) %>% 
  arrange(qAICc) %>%
  mutate('delta-qAICc' = qAICc - first(qAICc)) %>%
  gt() %>%
  fmt_number(columns=c(`qAICc`, `delta-qAICc`), decimals=2))

(qAICc_gt_tbl_Nsurv_quad <- tibble(
  Model = c("desert leaf + ant treatment + desert leaf*ant treatment + 
            desert leaf^2 + desert leaf^2*ant treatment",
            "desert plant + ant treatment + desert plant*ant treatment + 
            desert plant^2 + desert plant^2*ant treatment",
            "desert leaf + ants >2days + desert leaf*ants >2days + 
            desert leaf^2 + desert leaf^2*ants >2days",
            "desert plant + ants >2days + desert plant*ants >2days + 
            desert plant^2 + desert plant^2*ants >2days"),
  Response = c(rep("Nymph survivorship",4)),
  qAICc = c(
    get_qaicc(des_Nsurv.desL.ant_treatment.quad.fits),
    get_qaicc(des_Nsurv.desP.ant_treatment.quad.fits),
    get_qaicc(des_Nsurv.desL.ants_gt2d.quad.fits),
    get_qaicc(des_Nsurv.desP.ants_gt2d.quad.fits))
  ) %>% 
  arrange(qAICc) %>%
  mutate('delta-qAICc' = qAICc - first(qAICc)) %>%
  gt() %>%
  fmt_number(columns=c(`qAICc`, `delta-qAICc`), decimals=2))

(qAICc_gt_tbl_hatch_quad <- tibble(
  Model = c("desert leaf + ant treatment + desert leaf*ant treatment + 
            desert leaf^2 + desert leaf^2*ant treatment",
            "desert plant + ant treatment + desert plant*ant treatment + 
            desert plant^2 + desert plant^2*ant treatment",
            "desert leaf + ants >2days + desert leaf*ants >2days + 
            desert leaf^2 + desert leaf^2*ants >2days",
            "desert plant + ants >2days + desert plant*ants >2days + 
            desert plant^2 + desert plant^2*ants >2days"),
  Response = c(rep("Egg survivorship",4)),
  qAICc = c(
    get_qaicc(des_Esurv.desL.ant_treatment.quad.fits),
    get_qaicc(des_Esurv.desP.ant_treatment.quad.fits),
    get_qaicc(des_Esurv.desL.ants_gt2d.quad.fits),
    get_qaicc(des_Esurv.desP.ants_gt2d.quad.fits))
  ) %>% 
  arrange(qAICc) %>%
  mutate('delta-qAICc' = qAICc - first(qAICc)) %>%
  gt() %>%
  fmt_number(columns=c(`qAICc`, `delta-qAICc`), decimals=2))
```

## Stats interpretation
```{r}
# function to get fits and confint
ci_for_stats <- function(data_in, model){

  data <- eval(substitute(
    data_in %>%
      add_column(fit = predict(model, newdata = ., type = 'response'))
    ))

  ## grad the inverse link function
  ilink <- family(model)$linkinv

  ## add fit and se.fit on the **link** scale
  data_fit_link <- eval(substitute(
    data %>% add_column(
  setNames(as_tibble(predict(model, data, se.fit = TRUE)[1:2]),
           c('fit_link','se_link'))) %>%
## create the interval and backtransform
  mutate(fit_resp  = ilink(fit_link),
         conf_upr = ilink(fit_link + (2 * se_link)),
         conf_lwr = ilink(fit_link - (2 * se_link)))
    ))

  return(data_fit_link)
}

# function to get max fit, plus upper and lower ci
get_max_fit <- function(data_in, model){

  max_fit <- eval(substitute(ci_for_stats(data_in=data_in, 
             model) %>%
  group_by(ant_treatment) %>%
  mutate(max = max(fit)) %>% filter(fit==max) %>% 
    select(ant_treatment, des, max)))

  max_upr <- eval(substitute(ci_for_stats(data_in=data_in, 
             model) %>%
  group_by(ant_treatment) %>%
  mutate(max_upr = max(conf_upr)) %>% filter(conf_upr==max_upr) %>% 
    select(ant_treatment, des_upr=des, max_upr)))

  max_lwr <- eval(substitute(ci_for_stats(data_in=data_in, 
             model) %>%
  group_by(ant_treatment) %>%
  mutate(max_lwr = max(conf_lwr)) %>% filter(conf_lwr==max_lwr) %>% 
    select(ant_treatment, des_lwr=des, max_lwr)))

  data_out <- max_fit %>% left_join(max_upr) %>% left_join(max_lwr)
  return(data_out)
}

## Hatch
# at -20,0,10 etc.
ci_for_stats(data_in=
               tibble(ant_treatment=c(rep("Control",5), rep("Removed",5)),
                      des=rep(c(-20,10,0,10,20),2)), 
             des_hatch.desL.ant_treatment.quad.fits$mai.qb)

# where is the maximum?
get_max_fit(data_in=tibble(ant_treatment=c(rep("Control",10000), rep("Removed",10000)),
                           des=rep(seq(-20,20,length.out=10000),2)),
            model=des_hatch.desL.ant_treatment.quad.fits$mai.qb)

# Nsurv
# at -20,0,10 etc.
ci_for_stats(data_in=
               tibble(ant_treatment=c(rep("Control",5), rep("Removed",5)),
                      des=rep(c(-20,10,0,10,20),2)), 
             des_Nsurv.desL.ant_treatment.quad.fits$mai.qb)

# where is the maximum?
get_max_fit(data_in=tibble(ant_treatment=c(rep("Control",10000), rep("Removed",10000)),
                           des=rep(seq(-20,20,length.out=10000),2)),
            model=des_Nsurv.desL.ant_treatment.quad.fits$mai.qb)

# Esurv
# at -20,0,10 etc.
ci_for_stats(data_in=
               tibble(ant_treatment=c(rep("Control",5), rep("Removed",5)),
                      des=rep(c(-20,10,0,10,20),2)), 
             des_Esurv.desL.ant_treatment.quad.fits$mai.qb)

# where is the maximum?
get_max_fit(data_in=tibble(ant_treatment=c(rep("Control",10000), rep("Removed",10000)),
                           des=rep(seq(-20,20,length.out=10000),2)),
            model=des_Esurv.desL.ant_treatment.quad.fits$mai.qb)
```

# Figures for FPO

## Figures
### Both colors
```{r}
# hatching success
des_hatch.desL.ant_treatment.quad.fits <- fit_surv_quad(data_in=surv_reg_data, 
                              survNom=n_hatched, survDen=med_t5eggs,
                              tenure=des, treatment=ant_treatment)

des_hatch.desL.ant_treatment.quad.qb.ci <- 
  ci_for_plot(data_in=surv_reg_data, 
              tenure=des, treatment=ant_treatment,
              survNom=n_hatched, survDen=med_t5eggs,
              model = des_hatch.desL.ant_treatment.quad.fits$mai.qb)

des_hatch.desL.ant_treatment.quad.qb.p <- plot_ci(ci_data=des_hatch.desL.ant_treatment.quad.qb.ci, 
                                  tenure=des, treatment=ant_treatment,
             survNom = n_hatched, survDen = med_t5eggs) +
    xlab("Latest desertion date from leaf\nrelative to hatch") +
    ylab("Hatching success\n(nymphs/eggs)") + 
    scale_color_manual(values = c("brown1", "black")) +
    scale_fill_manual(values = c("brown1", "black"))

# nymph survivorship
des_Nsurv.desL.ant_treatment.quad.fits <- fit_surv_quad(data_in=surv_reg_data, 
                              survNom=n_survived, survDen=n_hatched,
                              tenure=des, treatment=ant_treatment)

des_Nsurv.desL.ant_treatment.quad.qb.ci <- 
  ci_for_plot(data_in=surv_reg_data, 
              tenure=des, treatment=ant_treatment,
              survNom=n_survived, survDen=n_hatched,
              model = des_Nsurv.desL.ant_treatment.quad.fits$mai.qb)

des_Nsurv.desL.ant_treatment.quad.qb.p <- plot_ci(ci_data=des_Nsurv.desL.ant_treatment.quad.qb.ci, 
                                  tenure=des, treatment=ant_treatment,
             survNom = n_survived, survDen = n_hatched) +
    xlab("Latest desertion date from leaf\nrelative to hatch") +
    ylab("Survivorship of nymphs\n(5th instars/nymphs)") + 
    scale_color_manual(values = c("brown1", "black")) +
    scale_fill_manual(values = c("brown1", "black"))

(FPO_desL.ant_treatment_allCols <- egg::ggarrange(des_hatch.desL.ant_treatment.quad.qb.p+
                 theme(legend.position="none"),
               des_Nsurv.desL.ant_treatment.quad.qb.p+
                 theme(legend.position="none")))

ggsave("FPO_desL.ant_treatment_allCols.png",
       FPO_desL.ant_treatment_allCols,
       device="png", width = 150, height = 200, units = "mm", dpi = "print")
```

### Black only
```{r}
# hatching success
des_hatch.desL.ant_treatment.quad.qb.p_blackOnly <- 
  ggplot(data = filter(des_hatch.desL.ant_treatment.quad.qb.ci,
                       ant_treatment=="Removed"), 
       aes(x = des, y = fit), 
           color = "black", fill = "black") +
    theme_publilia() +
    geom_point(aes(y = n_hatched/med_t5eggs), alpha = 0.25, size = 4, shape=16, key_glyph=draw_key_rect) + 
    geom_line(aes(y = fit), size = 1, key_glyph=draw_key_rect) +
    geom_ribbon(aes(ymin = conf_upr, ymax = conf_lwr),
                    alpha = 0.5, color = NA, key_glyph=draw_key_rect) +
      xlab("Latest desertion date from leaf\nrelative to hatch") +
    ylab("Hatching success\n(nymphs/eggs)")

# nymph survivorship
des_Nsurv.desL.ant_treatment.quad.qb.p_blackOnly <- 
  ggplot(data = filter(des_Nsurv.desL.ant_treatment.quad.qb.ci,
                       ant_treatment=="Removed"), 
       aes(x = des, y = fit), 
           color = "black", fill = "black") +
    theme_publilia() +
    geom_point(aes(y = n_survived/n_hatched), alpha = 0.25, size = 4, shape=16, key_glyph=draw_key_rect) + 
    geom_line(aes(y = fit), size = 1, key_glyph=draw_key_rect) +
    geom_ribbon(aes(ymin = conf_upr, ymax = conf_lwr),
                    alpha = 0.5, color = NA, key_glyph=draw_key_rect) +
  xlab("Latest desertion date from leaf\nrelative to hatch") +
    ylab("Survivorship of nymphs\n(5th instars/nymphs)")

(FPO_desL.ant_treatment_blackOnly <- egg::ggarrange(des_hatch.desL.ant_treatment.quad.qb.p_blackOnly+
                 theme(legend.position="none"),
               des_Nsurv.desL.ant_treatment.quad.qb.p_blackOnly+
                 theme(legend.position="none")))

ggsave("FPO_desL.ant_treatment_blackOnly.png",
       FPO_desL.ant_treatment_blackOnly,
       device="png", width = 150, height = 200, units = "mm", dpi = "print")
```

hatching success
```{r}
# hatching success
des_hatch.desL.ants_gt2d.quad.fits <- fit_surv_quad(data_in=surv_reg_data, 
                              survNom=n_hatched, survDen=med_t5eggs,
                              tenure=des, treatment=ants_gt2d)

des_hatch.desL.ants_gt2d.quad.qb.ci <- 
  ci_for_plot(data_in=surv_reg_data, 
              tenure=des, treatment=ants_gt2d,
              survNom=n_hatched, survDen=med_t5eggs,
              model = des_hatch.desL.ants_gt2d.quad.fits$mai.qb)

des_hatch.desL.ants_gt2d.quad.qb.p <- plot_ci(ci_data=des_hatch.desL.ants_gt2d.quad.qb.ci, 
                                  tenure=des, treatment=ants_gt2d,
             survNom = n_hatched, survDen = med_t5eggs) +
    xlab("Latest desertion date from leaf\nrelative to hatch") +
    ylab("Hatching success\n(nymphs/eggs)") + 
    scale_color_manual(values = c("brown1", "black")) +
    scale_fill_manual(values = c("brown1", "black"))

plot_ci(ci_data=filter(des_hatch.desL.ants_gt2d.quad.qb.ci, 
                       ants_gt2d ), 
                                  tenure=des, treatment=ants_gt2d,
             survNom = n_hatched, survDen = med_t5eggs) +
    xlab("Latest desertion date from leaf\nrelative to hatch") +
    ylab("Hatching success\n(nymphs/eggs)") + 
    scale_color_manual(values = c("brown1", "black")) +
    scale_fill_manual(values = c("brown1", "black"))

```

nymph survivorship
```{r}
# nymph survivorship
des_Nsurv.desL.ants_gt2d.quad.fits <- fit_surv_quad(data_in=surv_reg_data, 
                              survNom=n_survived, survDen=n_hatched,
                              tenure=des, treatment=ants_gt2d)

des_Nsurv.desL.ants_gt2d.quad.qb.ci <- 
  ci_for_plot(data_in=surv_reg_data, 
              tenure=des, treatment=ants_gt2d,
              survNom=n_survived, survDen=n_hatched,
              model = des_Nsurv.desL.ants_gt2d.quad.fits$mai.qb)

des_Nsurv.desL.ants_gt2d.quad.qb.p <- plot_ci(ci_data=des_Nsurv.desL.ants_gt2d.quad.qb.ci, 
                                  tenure=des, treatment=ants_gt2d,
             survNom = n_survived, survDen = n_hatched) +
    xlab("Latest desertion date from leaf\nrelative to hatch") +
    ylab("Survivorship of nymphs\n(5th instars/nymphs)") + 
    scale_color_manual(values = c("brown1", "black")) +
    scale_fill_manual(values = c("brown1", "black"))
```