### 2023-08-26 scripts/Publilia-06-fitness_sensu_Zink.Rmd

---
title: '06 - PubliliaFitness - Estimating lifetime fecundity and fitness using formulas from Zink 2003'
author: "Micah Fletcher"
date: '`r Sys.Date()`'
output: 
  rmdformats::robobook:
  fig_width: 9
toc_depth: 3
---

# Load libraries

```{r}
library(multcompView)
source(here("R/Publilia-05-fitness-functions.R"))
```

# 0 - Data import

```{r}
surv_reg_data <- read_rds(here("saved_rds/surv_reg_data.rds"))
```

## Comparing to Zink 2003 method

### 3 Zink 2003 model:
Lifetime fitness (number of eggs that hatched):
w(t_i) = sum(from i=1, to i=N, P_i*g(t_i))
P = number of eggs invested in a clutch (all the eggs go to the initiator?)
t_i = guarding duration in clutch i
N = lifetime number of clutches
g(t) = relationship between guarding duration and hatching success
 (assume g(t) equivalent across all clutches 1 to N)

* I think the main reason Zink found equivalent fitness is that he attributed all the eggs in a clutch to each primary female. Thus an early deserter is given all the eggs from a clutch, even if she leaves within a few days and another female arrives and lays additional eggs. The final nail in the coffin will be to do the exact same "zink" estimates but this time only attributing the egg to each female proportional to the time she spent on the clutch relative to other females.
```{r}
surv_reg_data_zink <- focalMb4h_sc2h %>%
  filter(!is.na(initiation_date), !is.na(guard_duration), !ants_removed) %>%
  filter(mom_at_init) %>% # primary females
  group_by(location) %>%
    summarize(max_primary_dur = max(guard_duration, na.rm=T),
              females_removed = first(females_removed),
              initiation_date = first(initiation_date)) %>%
  ungroup() %>%
  left_join(top5eggs_summary) %>%
  select(location, max_primary_dur,
         females_removed, initiation_date, med_t5eggs) %>%
  left_join(select(clutch_unambig_NOr_revised_ENAa_summary_imfilt_success_summary_v2,
                   -c(ants_removed, females_removed, initiation_date, n_eggs)),
            by=c("location"="NOr_rev_loc")) %>%
  select(location, initiation_date, hatch_date, 
         max_primary_dur, med_t5eggs, n_hatched, n_survived,
         females_removed)

ggplot(data=surv_reg_data_zink, 
       aes(x=max_primary_dur, y=n_hatched/med_t5eggs)) +
  geom_point() +
  geom_smooth(method=lm)

ggplot(data=filter(surv_reg_data_zink, initiation_date<=as_date("2019-06-10")), 
       aes(x=max_primary_dur, y=n_hatched/med_t5eggs)) +
  geom_point() +
  geom_smooth(method=lm)

ggplot(data=filter(surv_reg_data_zink, initiation_date<=as_date("2019-06-10"), !females_removed), 
       aes(x=max_primary_dur, y=n_hatched/med_t5eggs)) +
  geom_point() +
  geom_smooth(method=lm)

zink_hatch.m <- surv_reg_data_zink %>%
  filter(initiation_date<=as_date("2019-06-10"), !is.na(n_hatched/med_t5eggs)) %>%
lm(n_hatched/med_t5eggs ~ max_primary_dur, data=.)

summary(zink_hatch.m) # use this for predicting hatching success

zink_hatch.m.norem <- surv_reg_data_zink %>%
  filter(initiation_date<=as_date("2019-06-10"), !is.na(n_hatched/med_t5eggs), !females_removed) %>%
lm(n_hatched/med_t5eggs ~ max_primary_dur, data=.)

summary(zink_hatch.m.norem)

est_fitness_clutch_zink <- focalMb4h_sc2h %>%
  filter(!is.na(initiation_date), !ants_removed) %>%
  group_by(location) %>%
    mutate(max_primary_dur = max(mom_at_init*guard_duration, na.rm=T)) %>% # non-primary females are not counted
  ungroup() %>%
  left_join(top5eggs_summary) %>%
  left_join(clutch_momDays_b4h) %>%
  select(Female_ID, location, guard_duration, max_primary_dur, desert_rel2h=est_rel2h, focalMb4h_sc2h, mDb4h_sc2h,
         females_removed, initiation_date, med_t5eggs, 
         mom_at_init, clutch_visited_order, clutch_initiated_order) %>%
  left_join(select(clutch_unambig_NOr_revised_ENAa_summary_imfilt_success_summary_v2,
                   -c(ants_removed, females_removed, initiation_date, n_eggs)),
            by=c("location"="NOr_rev_loc")) %>%
  select(Female_ID, location, initiation_date, hatch_date, 
         guard_duration, max_primary_dur, desert_rel2h, focalMb4h_sc2h, mDb4h_sc2h,
         med_t5eggs, n_hatched, n_survived, females_removed,
         mom_at_init, clutch_visited_order, clutch_initiated_order) %>%
  mutate(propD_b4h = focalMb4h_sc2h/mDb4h_sc2h,
         est_eggs_clutch = med_t5eggs*propD_b4h,
         zink_est_hatch_clutch_v1 = med_t5eggs*(coef(zink_hatch.m)[[1]]+guard_duration*coef(zink_hatch.m)[[2]]),
         zink_est_hatch_clutch_v2 = est_eggs_clutch*(coef(zink_hatch.m)[[1]]+guard_duration*coef(zink_hatch.m)[[2]])) %>%
  group_by(Female_ID) %>%
    mutate(zink_est_hatch_life_v1 = sum(zink_est_hatch_clutch_v1, na.rm=T),
           zink_est_hatch_life_v2 = sum(zink_est_hatch_clutch_v2, na.rm=T)) %>%
  ungroup()

est_fitness_life_strats_Zink <- est_fitness_clutch_zink %>%
  group_by(Female_ID) %>%
  # clutch visited order == 0 means she visited before initiation, so it shouldn't count
  mutate(started_by =
           ifelse(any(clutch_visited_order==1&!is.na(clutch_initiated_order)&clutch_initiated_order==1),
                  "initiating",
                  ifelse(any(clutch_visited_order==1&is.na(clutch_initiated_order)|
                             clutch_visited_order==1&!is.na(clutch_initiated_order)&clutch_initiated_order!=1),
                         "visiting", "other")),
         n_clutches_with_eggs_iv = sum(!is.na(clutch_visited_order)&(clutch_visited_order>0), na.rm=T)) %>%
  ungroup() %>%
  filter(clutch_visited_order==1) %>%
  # mutate(desert_cat = ifelse(started_by=="initiating",
  #                                ifelse(desert_rel2h <= -16, "desert",
  #                                       ifelse(desert_rel2h > -16 & desert_rel2h <= 4, "eggs", 
  #                                              ifelse(desert_rel2h >= 4,"nymphs",
  #                                                     NA))),
  #                                ifelse(started_by=="visiting", "visit", 
  #                                       NA)),
  mutate(desert_cat = ifelse(started_by=="initiating",
                                 ifelse(guard_duration <= 10, "desert",
                                        ifelse(guard_duration > 10 & desert_rel2h <= 4, "eggs", 
                                               ifelse(guard_duration >= 4,"nymphs",
                                                      NA))),
                                 ifelse(started_by=="visiting", "visit", 
                                        NA)),
         zink_dur_cat = ifelse(started_by=="initiating",
                               ifelse(guard_duration<12, "desert",
                                      ifelse(guard_duration>=12, "stay", NA)),
                               ifelse(started_by=="visiting", "visit", 
                                        NA))) %>%
  distinct(Female_ID, females_removed,
           guard_duration, max_primary_dur,
           started_by, desert_cat, zink_dur_cat,
           zink_est_hatch_clutch_v1, zink_est_hatch_clutch_v2,
           zink_est_hatch_life_v1, zink_est_hatch_life_v2,
           n_clutches_with_eggs_iv)

# v1: Zink 2003 attributes all egg in a clutch to the primary female, even if they desert early and a visitor laid additional eggs
est_fitness_life_strats_Zink %>%
  filter(!females_removed) %>%
  ggplot(data=., aes(x=zink_dur_cat, y=zink_est_hatch_life_v1, fill=zink_dur_cat)) +
  geom_boxplot(size=1.5) +
  geom_jitter(pch = 21, size=3, width=0.35) +
  stat_summary(fun=mean, geom="point", shape=4, size=6) +
  ggtitle("v1: all eggs in clutch to primary female")

# v2: all calculations the same, but give eggs to females in proportion to the time they spent at clutch relative to other females
est_fitness_life_strats_Zink %>%
  filter(!females_removed) %>%
  ggplot(data=., aes(x=zink_dur_cat, y=zink_est_hatch_life_v2, fill=zink_dur_cat)) +
  geom_boxplot(size=1.5) +
  geom_jitter(pch = 21, size=3, width=0.35) +
  stat_summary(fun=mean, geom="point", shape=4, size=6) +
  ggtitle("v2: eggs to female by proportion of time guarding")

# same but separating females that laid one vs. multiple clutches
est_fitness_life_strats_Zink %>%
  filter(!females_removed) %>%
  ggplot(data=., aes(x=zink_dur_cat, y=zink_est_hatch_life_v1, fill=zink_dur_cat)) +
  geom_boxplot(size=1.5) +
  geom_jitter(pch = 21, size=3, width=0.35) +
  stat_summary(fun=mean, geom="point", shape=4, size=6) +
  facet_wrap(vars(n_clutches_with_eggs_iv!=1)) +
  ggtitle("v1: all eggs in clutch to primary female")

est_fitness_life_strats_Zink %>%
  filter(!females_removed) %>%
  ggplot(data=., aes(x=zink_dur_cat, y=zink_est_hatch_life_v2, fill=zink_dur_cat)) +
  geom_boxplot(size=1.5) +
  geom_jitter(pch = 21, size=3, width=0.35) +
  stat_summary(fun=mean, geom="point", shape=4, size=6) +
  facet_wrap(vars(n_clutches_with_eggs_iv!=1)) +
  ggtitle("v2: eggs to female by proportion of time guarding")

est_fitness_life_strats_Zink %>%
  filter(!females_removed) %>%
  ggplot(data=., aes(x=desert_cat, y=zink_est_hatch_life_v1, fill=desert_cat)) +
  geom_boxplot(size=1.5) +
  geom_jitter(pch = 21, size=3, width=0.35) +
  stat_summary(fun=mean, geom="point", shape=4, size=6) +
  ggtitle("v1: all eggs in clutch to primary female")

# same but using desert_cat instead of Zink 2003 cats
est_fitness_life_strats_Zink %>%
  filter(!females_removed) %>%
  ggplot(data=., aes(x=desert_cat, y=zink_est_hatch_life_v1, fill=desert_cat)) +
  geom_boxplot(size=1.5) +
  geom_jitter(pch = 21, size=3, width=0.35) +
  stat_summary(fun=mean, geom="point", shape=4, size=6) +
  facet_wrap(vars(n_clutches_with_eggs_iv!=1)) +
  ggtitle("v1: all eggs in clutch to primary female")

est_fitness_life_strats_Zink %>%
  filter(!females_removed) %>%
  ggplot(data=., aes(x=desert_cat, y=zink_est_hatch_life_v2, fill=desert_cat)) +
  geom_boxplot(size=1.5) +
  geom_jitter(pch = 21, size=3, width=0.35) +
  stat_summary(fun=mean, geom="point", shape=4, size=6) +
  ggtitle("v2: eggs to female by proportion of time guarding")

est_fitness_life_strats_Zink %>%
  filter(!females_removed) %>%
  ggplot(data=., aes(x=desert_cat, y=zink_est_hatch_life_v2, fill=desert_cat)) +
  geom_boxplot(size=1.5) +
  geom_jitter(pch = 21, size=3, width=0.35) +
  stat_summary(fun=mean, geom="point", shape=4, size=6) +
  facet_wrap(vars(n_clutches_with_eggs_iv!=1)) +
  ggtitle("v2: eggs to female by proportion of time guarding")

```

## Comparing to Zink 2003 fitness measures
Lifetime fitness (number of eggs that hatched):
w(t_i) = sum(from i=1, to i=N, P_i*g(t_i))
P = number of eggs invested in a clutch (all the eggs go to the initiator?)
t_i = guarding duration in clutch i
N = lifetime number of clutches
g(t) = relationship between guarding duration and hatching success
 (assume g(t) equivalent across all clutches 1 to N)

* The main reason Zink found equivalent fitness is that he attributed all the eggs in a clutch to each primary female. Thus an early deserter is given all the eggs from a clutch, even if she leaves within a few days and another female arrives and lays additional eggs.
```{r}
ggplot(data=filter(surv_reg_data_zink, initiation_date<=as_date("2019-06-10")), 
       aes(x=max_primary_dur, y=n_hatched/med_t5eggs)) +
  geom_point() +
  geom_smooth(method=lm)

zink_hatch.m <- surv_reg_data_zink %>%
  filter(initiation_date<=as_date("2019-06-10"), !is.na(n_hatched/med_t5eggs)) %>%
lm(n_hatched/med_t5eggs ~ max_primary_dur, data=.)

summary(zink_hatch.m) # use this for predicting hatching success
```

```{r}
zink_fitness_plot_dunn <- function(data, y, group, multcompLetters, letterSize, ymax=160){
  (p <- eval(substitute(data %>%
    filter(!females_removed, !is.na(desert_cat)) %>%
    ggplot(data=., aes(x=group, y=y)) + ylim(-ymax*0.05,ymax) +
    geom_boxplot(aes(fill=group), 
                 position = position_dodge(0.75), size = 1.5, key_glyph=draw_key_rect) +
    stat_summary(aes(fill=group), 
                 fun=mean, geom="point", shape=4, size=3, color="black", key_glyph=draw_key_rect) +
    #ylab("Lifetime fecundity\n(eggs laid)")  +
    xlab("Guarding strategy for first clutch") +
    theme_publilia() +
    geom_text(data=multcompLetters, aes(x=x, y=-ymax*0.05, label=label), fontface="bold", size=letterSize))))
  return(p)
}
```

```{r}
# v1: Zink 2003 attributes all egg in a clutch to the primary female, even if they desert early and a visitor laid additional eggs
zink_fitness_stats_v1 <- 
  fitness_KW_test(data=est_fitness_life_strats_Zink, y=zink_est_hatch_life_v1, group=zink_dur_cat)
(Zink_fitness_v1.p <- 
  zink_fitness_plot_dunn(data=est_fitness_life_strats_Zink, y=zink_est_hatch_life_v1, group=zink_dur_cat, 
                    multcompLetters=zink_fitness_stats_v1$multcompLetters,
                    letterSize=4, ymax=150) +
  ylab("Lifetime fitness: Zink 2003\n(all eggs to primary mother,\nlinear model predicts hatching)") +
    # geom_jitter(aes(fill=ants_removed), pch = 21, size=3, width=0.35) + 
    theme(legend.position="none",
          axis.text.x=element_text(angle=45, vjust=1, hjust=1)))

est_fitness_life_strats_Zink %>%
  filter(!females_removed, !is.na(desert_cat)) %>%
  group_by(zink_dur_cat) %>% 
    summarize(mean_fitness=mean(zink_est_hatch_life_v1),
              median_fitness=median(zink_est_hatch_life_v1))

# v2: all calculations the same, but give eggs to females in proportion to the time they spent at clutch relative to other females
zink_fitness_stats_v2 <- 
  fitness_KW_test(data=est_fitness_life_strats_Zink, y=zink_est_hatch_life_v2, group=zink_dur_cat)
(Zink_fitness_v2.p <- 
  zink_fitness_plot_dunn(data=est_fitness_life_strats_Zink, y=zink_est_hatch_life_v2, group=zink_dur_cat, 
                    multcompLetters=zink_fitness_stats_v2$multcompLetters,
                    letterSize=4, ymax=150) +
  ylab("Lifetime fitness: modified from Zink 2003\n(eggs proportional to guarding,\nlinear model predicts hatching)") +
    # geom_jitter(aes(fill=ants_removed), pch = 21, size=3, width=0.35) + 
    theme(legend.position="none",
          axis.text.x=element_text(angle=45, vjust=1, hjust=1)))

(Zink_fitness.p.final <- egg::ggarrange(
  Zink_fitness_v1.p, Zink_fitness_v2.p, 
  labels=c("A","B"),
  ncol=2, nrow=1))

ggsave("Zink_fitness.final.png",
       Zink_fitness.p.final,
       device = "png",
       width = 200, height = 120, units = "mm", dpi = "print")

# same but separating females that laid one vs. multiple clutches
zink_fitness_data.semel <- est_fitness_life_strats_Zink %>% 
  filter(n_clutches_with_eggs_iv==1)
zink_fitness_stats_v1.semel <- 
  fitness_KW_test(data=zink_fitness_data.semel, y=zink_est_hatch_life_v1, group=zink_dur_cat)
(Zink_fitness_v1.semel.p <- 
  zink_fitness_plot_dunn(data=zink_fitness_data.semel, y=zink_est_hatch_life_v1, group=zink_dur_cat, 
                    multcompLetters=zink_fitness_stats_v1.semel$multcompLetters,
                    letterSize=4, ymax=150) +
  ylab("Lifetime fitness: Zink 2003\n(all eggs to primary mother,\nlinear model predicts hatching)") +
    ggtitle("Single-clutch mothers") +
    # geom_jitter(aes(fill=ants_removed), pch = 21, size=3, width=0.35) + 
    theme(legend.position="none",
          axis.title.x = element_blank(),axis.text.x=element_blank()))

zink_fitness_data.itero <- est_fitness_life_strats_Zink %>% 
  filter(n_clutches_with_eggs_iv>1)
zink_fitness_stats_v1.itero <- 
  fitness_KW_test(data=zink_fitness_data.itero, y=zink_est_hatch_life_v1, group=zink_dur_cat)
(Zink_fitness_v1.itero.p <- 
  zink_fitness_plot_dunn(data=zink_fitness_data.itero, y=zink_est_hatch_life_v1, group=zink_dur_cat, 
                    multcompLetters=zink_fitness_stats_v1.itero$multcompLetters,
                    letterSize=4, ymax=150) + ggtitle("Multi-clutch mothers") +
    # geom_jitter(aes(fill=ants_removed), pch = 21, size=3, width=0.35) + 
    theme(legend.position="none",
          axis.title.x = element_blank(),axis.text.x=element_blank(),
          axis.title.y = element_blank(),axis.text.y=element_blank()))

zink_fitness_data.semel <- est_fitness_life_strats_Zink %>% 
  filter(n_clutches_with_eggs_iv==1)
zink_fitness_stats_v2.semel <- 
  fitness_KW_test(data=zink_fitness_data.semel, y=zink_est_hatch_life_v2, group=zink_dur_cat)
(Zink_fitness_v2.semel.p <- 
  zink_fitness_plot_dunn(data=zink_fitness_data.semel, y=zink_est_hatch_life_v2, group=zink_dur_cat, 
                    multcompLetters=zink_fitness_stats_v2.semel$multcompLetters,
                    letterSize=4, ymax=150) +
  ylab("Lifetime fitness: modified from Zink 2003\n(all eggs to primary mother,\nlinear model predicts hatching)") +
    # geom_jitter(aes(fill=ants_removed), pch = 21, size=3, width=0.35) + 
    theme(legend.position="none",
          axis.text.x=element_text(angle=45, vjust=1, hjust=1)))

zink_fitness_data.itero <- est_fitness_life_strats_Zink %>% 
  filter(n_clutches_with_eggs_iv>1)
zink_fitness_stats_v2.itero <- 
  fitness_KW_test(data=zink_fitness_data.itero, y=zink_est_hatch_life_v2, group=zink_dur_cat)
(Zink_fitness_v2.itero.p <- 
  zink_fitness_plot_dunn(data=zink_fitness_data.itero, y=zink_est_hatch_life_v2, group=zink_dur_cat, 
                    multcompLetters=zink_fitness_stats_v2.itero$multcompLetters,
                    letterSize=4, ymax=150) +
    # geom_jitter(aes(fill=ants_removed), pch = 21, size=3, width=0.35) + 
    theme(legend.position="none",
          axis.text.x=element_text(angle=45, vjust=1, hjust=1),
          axis.title.y = element_blank(),axis.text.y=element_blank()))

(Zink_fitness.parity.p.final <- egg::ggarrange(
  Zink_fitness_v1.semel.p, Zink_fitness_v1.itero.p,
  Zink_fitness_v2.semel.p, Zink_fitness_v2.itero.p,
  labels=c("A","B","C","D"),
  ncol=2, nrow=2))

ggsave("Zink_fitness.parity.final.png",
       Zink_fitness.parity.p.final,
       device = "png",
       width = 200, height = 200, units = "mm", dpi = "print")

```

why such a small first clutch and larger future clutches for early deserters? early deserters lay more or larger clutches?
```{r}
est_fitness_life_strats_Zink %>%
  filter(!females_removed) %>%
  mutate(parity=ifelse(n_clutches_with_eggs_iv==1,"semel","itero")) %>%
  count(parity,zink_dur_cat)

est_fitness_life_strats_Zink %>%
  filter(!females_removed) %>%
  mutate(parity=ifelse(n_clutches_with_eggs_iv==1,"semel","itero")) %>%
  group_by(zink_dur_cat,parity) %>%
  summarize(n=n(),
            mean_n_clutches=mean(n_clutches_with_eggs_iv),
            mean_max_primary_dur=mean(max_primary_dur)) %>%
  ungroup()

est_fitness_life_strats_Zink %>%
  filter(!females_removed) %>%
  mutate(parity=ifelse(n_clutches_with_eggs_iv==1,"semel","itero")) %>%
  group_by(desert_cat,parity) %>%
  summarize(n=n(),
            mean_n_clutches=mean(n_clutches_with_eggs_iv),
            mean_max_primary_dur=mean(max_primary_dur)) %>%
  ungroup()

fitness_life_strats_data %>%
  filter(!females_removed) %>%
  mutate(parity=ifelse(n_clutches_with_eggs_iv==1,"semel","itero")) %>%
  group_by(parity, ants_removed, desert_cat) %>%
  summarize(n=n(),
            mean_n_clutches=mean(n_clutches_with_eggs_iv)) %>%
  ungroup()

```